Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-05-12 rmarkdown/paper.Rnw
new paper version

The strand cross-correlation was proposed by Kharchenko et al., 2008
\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
metrics. The SCC curve is defined as:

@@

is a re-parametrization of the following relationship:

@@

with $\beta_1 = 1 / \gamma$ and $\beta_2 = - \kappa / \gamma$. In this
setting, $\gamma$ can be considered as the large - depth
$\mbox{URCR}$, i.e. the ratio between the number of unique positions
and depth as the depth limits to infinity.

On the other hand, to interpret $\beta_2$, if we assume that the
$\gamma$ value is already given, then by expressing $\kappa$ as a
function of $\mbox{ARC}$ and $\mbox{URCR}$ we can observe the
following identities:

@@

where $\mbox{npos}$ and $\mbox{width}$ denote the region's number of
unique position where the reads are aligned to and the region's width
respectively. This is particularly important, $\gamma$ approximates
the true $\mbox{URCR}$ as the sequencing depth increases, which
implies that $- \kappa / \gamma$ can be interpreted as the large
sequencing depth bias of the $\mbox{ARC}$ since as the depth
increases, the first term of $\kappa / \gamma$ is going to approximate
the average read coverage:

@@

Hence, we define the $\mbox{ARC}$ bias as $- \kappa / \gamma$.

@@

We compared the estimated binding events predicted by the MOSAiCS +
dPeak pipeline using reads generated by ChIP-exo, PE and SE ChIP-Seq
protocols. We called peaks at a $5\%$ FDR level using MOSAiCS (the
$\mbox{GC content} + \mbox{Mappability}$ model for ChIP-exo and the
$\mbox{Input only}$ model for PE and SE ChIP-Seq). Then, we
deconvolved the peak into binding events with dPeak (version 2.0.1) by
considering a maximum of 5 binding sites on each peak. To avoid false
positives we only considered ChIP-exo peaks with average ChIP counts
greater than 3000 that overlapped both the SE and PE ChIP-Seq peaks,
we considered other cutoff values but still obtained results similar
to what we presented in this paper.

@@

 % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliographystyle{vancouver}
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}

>>>>

The strand cross-correlation was proposed by Kharchenko et al.,
2008\nocite{strandcc} and it may be one of the most used of the
ChIP-Seq QC metrics. The SCC curve is defined as:

@@

where $\mbox{npos}$ and $\mbox{width}$ denote the region's number of
unique position where the reads are aligned to and the region's width
respectively. Is a re-parametrization of the following relationship:

@@

with $\beta_1 = 1 / \gamma$ and $\beta_2 = - \kappa / \gamma$. In this
setting, $\gamma$ can be considered as the large - depth
$\mbox{URCR}$, i.e. the ratio between the number of unique positions
and depth as the depth limits to infinity. On the other hand, to
interpret $\beta_2 = - \kappa / \gamma $, by expressing $\kappa$ as a
function of $\mbox{ARC}$ and $\mbox{URCR}$ and assuming that $\gamma$
is already estimated, we can observe the following identities:

@@

This is important: $\gamma$ approximates the $\mbox{URCR}$ as the
sequencing depth increases, which implies that $- \kappa / \gamma$ can
be interpreted as the large sequencing depth bias of the $\mbox{ARC}$
since as the depth increases, the first term of $\kappa / \gamma$ is
going to approximate the average read coverage:

@@

Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the
$\mbox{ARC}$ bias as the sequencing depth increases.

@@

We compared the estimated binding events predicted by the
\textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
ChIP-exo, PE and SE ChIP-Seq protocols. We called peaks at a $5\%$ FDR
level using \textbf{MOSAiCS} (the $\mbox{GC content} +
\mbox{Mappability}$ model for ChIP-exo and the $\mbox{Input only}$
model for PE and SE ChIP-Seq). Then, we deconvolved the peak into
binding events with \textbf{dPeak} (version 2.0.1) by considering a
maximum of 5 binding sites on each peak. To avoid false positives we
only considered ChIP-exo peaks with average ChIP counts greater than
3000 that overlapped both the SE and PE ChIP-Seq peaks, we considered
other cutoff values but still obtained results similar to what we
presented in this paper.

@@

 % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliographystyle{vancouver}
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}
\nocite{meme}
